Apropos Replacement \newline (Extended Abstract)

### 2011-12-25 paper.tex
More conversion from active to passive.

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the keywords defined in the NAME section of the man
pages, the keywords used throughout the body of the page also play a role.
For example queries like \textit{``installing new software package''}
will now get results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This allows us to support more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

@@

In this implementation, both word stemming and spelling correction in our implementation are used to make the search experience more smooth.
\end{description}
\begin{description}
\item[Bookkeeping of man page metadata] \hfill \\
In this implementation we have gone ahead and also indexed additional metadata
related to the man pages, for example we are indexing an md5 hash of the man
page, the device id, inode number and modification time of the man page file so
as to support incremental updating of the index as new man pages are installed
or the old ones are modified/updated.

Similarly we are also maintaining a separate index of all the aliases of a man
page. This will allow us to get rid of all the hard or symbolic links of the man
pages which are scattered throughout the filesystem, and also help in getting rid
of the MKLINK flags in the makefiles.
\end{description}
\subsection{Tools Used}
The two main tasks that we needed to perform were parsing of the man pages and
building an index of the data, and to do these tasks we have used \textit{libmandoc} [6]
and \textit{sqlite} [7] in our implementation.
\begin{description}
\item[\textit{libmandoc}] \hfill \\
\textit{libmandoc} is a library interface to a validating compiler. It provides
interface to parse and build an AST (Abstract Syntax Tree) of the man page, 
and interface for traversing that tree in order to extract the data from nodes
which are of our interest.
\end{description}
\begin{description}
\item[\textit{sqlite}] \hfill \\
\textit{sqlite} is an embedded relational database management system providing a
relatively small and easy to use C library interface. One of the main reason
for choosing it over the myriad of other possible options was that it provides
in built support for full text search through it's FTS virtual table module [8].
The FTS module can be accessed using pretty much standard SQL syntax, and it is
still flexible enough to allow us to implement our own ranking function as
per our needs. Besides that we have been able to utilize the RDBMS support of
\textit{Sqlite} to store additional metadata in the form of normal database
tables without any hassles.
\end{description}
\section{Implementation Details}
Due to space constraints we cannot go into enough implementation details in this
abstract but we would like to talk a bit about some important components of this
project:
\begin{description}
\item[makemandb(1)] \hfill \\
\textit{makemandb(1)} [9] is the key component of this implementation. It
traverses the filesystem, reads the raw man page source files, feeds them to them
\textit{libmandoc} and then stores the extracted data in the database using
\textit{sqlite}.
\end{description}
\begin{description}
\item[apropos(1)] \hfill \\
This is the version of \textit{apropos(1)} written from scratch utilizing the
full text search functionality of \textit{sqlite}. Unlike the classical version
of \textit{apropos(1)} it only displays the top 10 results relevant to the user
query and most of the times this is sufficient as we shall see later. Also, it
shows a brief snippet for each of the search results, making it more easy to
find out the relevant documents.
\end{description}
\begin{description}
\item[apropos-utils(3)] \hfill \\
\textit{apropos-utils(3)} [10] is a small library interface provided with this
implementation.
It provides functions for querying the FTS index and processing the results in
a user supplied callback function. It's main purpose is to develop different
interfaces on top of this implementation for different use cases. For example we
have built a small CGI application using it for doing the searches from a web
browser, similarly we also built an IRC bot using this 
\end{description}
\section{Results}
We would like to show how our implementation solves the problems cited earlier
with the classical version of \textit{apropos(1)}.
\footnote{Although this implementation of \textit{apropos(1)} returns 10 results
in normal case for a query but in the following listings we have snipped the
results to save space
}
\lstset{language = sh, caption={Add new user},}
\begin{lstlisting}
$ apropos ``add new user''
ssh-add(1)      adds private key identities to the authentication agent
...on the command line. If any file requires a passphrase, ssh-add
asks for the passphrase from the user. The passphrase is read from the
user's tty. ssh-add retries the last passphrase if multiple identity
files are given...

@@

mkdir(2)        make a directory file
...will contain the directory has been exhausted. EDQUOT The user's
quota of inodes on the file system on which the directory is being
created has been exhausted. EIO An I/O error occurred while making the
directory entry or...
\end{lstlisting}
\lstset{language = sh, caption={get process status},}
\begin{lstlisting}
$ apropos ``get process status''
ps(1)   process status
process status

netstat(1)      show network status
...use the fstat(1) command to find out which process or processes
hold references to a socket. The interface display...show network
status

bounce(8)       Postfix delivery status reports
...the master(8) process manager.  The bounce(8) daemon
processes two types of service requests: Append a recipient
(non-)delivery status record to a per message log file. Enqueue a
delivery status notification message, with a...

ocsp(1) Online Certificate Status Protocol utility
Online Certificate Status Protocol utility...it is in client mode. The
request(s) the responder processes can be either specified on the
command line (using...

fstat(1)        display status of open files
...the owner of the process (effective UID). CMD The command name of
the process. PID The process ID. FD The...display status of open files
\end{lstlisting}

\begin{figure}[htp]
\includegraphics[scale=0.30]{/home/abhinav/development/AsiaBSDCon/spell.png}
\caption{The spell corrector and the web interface in action}
\label{}
\end{figure}
\section{Related Work}
There are at least two projects which are related to this in some way.
\begin{description}
\item[man-db] \hfill \\
\textit{man-db} [11] is a complete implementation of the man page documentation
system and it is used on a number of GNU/Linux distributions.\textit{man-db}
takes an interesting approach for indexing the man page data. Unlike the
classical \textit{apropos(1)}, it uses a Berkley DB database but still it's index
is limited to the NAME section only. It adds an option `K' to man(1) to allow a
crude full text search but it is not very efficient nor effective.
\end{description}
\begin{description}
\item[mandocdb] \hfill \\
\textit{mandocdb} [12] is more in line with the goals of our project but it takes
a novel approach. It indexes the keywords extracted from the man pages in a
key-value store using btree(3) [13]. It also comes with it's own implementation
of \textit{apropos(1)} which performs search using this key-value store. The
main key point of this implementation is that it exploits the semantic structure
of the man pages. 
\end{description}
\section{Future Work}
\begin{description}
\item[Work on Ranking Algorithm] \hfill \\
We have used a ranking algorithm based on probabilistic model [14] of information
retrieval. It is essentially based on the Okapi BM25F algorithm [15] and uses
certain parameters whose values are usually dependent on the corpus and the
search application. For example we have used certain weights for different
sections of man pages. At the moment these values have been determined manually
but we would like to use some machine learning techniques in the coming days to
automate this.
\end{description}
\begin{description}
\item[Fix Some Loose Ends In Parsing of man pages] \hfill \\
Although the parsing routine is working fine for most purposes on the man pages
supplied with NetBSD and also on most of the man pages found in pkgsrc. But there
are perhaps a few corner cases still to be fixed. Besides that another issue that
is to be addressed is the parsing of the escape sequences which are used
extensively in the \textit{man(7)} [16] based man pages. We would like to address
these issues in the coming days.
\end{description}
\section{Acknowledgement}
This project was developed as part of \textbf{Google Summer of Code 2011} [17],
so thanks to Google for sponsoring it. We would also like to thank
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. We would also like to thank \textbf{David Young} who was involved with
this project closely and offered useful help and guidance. A special thanks goes
to \textbf{Thomas Klausner} who helped in writing and reviewing our man pages.
Thanks to Petra Zeidler for administering the GSoC program for \textit{The NetBSD
Foundation}.
\begin{thebibliography}{20}
\bibitem{1}
\emph{NetBSD manual page for apropos(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?apropos++NetBSD-5.1

>>>>

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the keywords defined in the NAME section of the man
pages, the keywords used throughout the body of the page also play a role.
For example queries like \textit{``installing new software package''}
will now get results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This enables support for more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

@@

In this implementation, both word stemming and spelling correction in our implementation are used to make the search experience more smooth.
\end{description}
\begin{description}
\item[Bookkeeping of man page metadata] \hfill \\
In this implementation additional metadata related to the man pages, for example
an md5 hash, the device id, inode number and modification time
of the man page file etc. are indexed and stored so
as to support fast and hassle free updating of the index as new man pages are
installed or the old ones are modified/updated.

Similarly a separate index of all the man page aliases is stored and maintained.
This provides an option to get rid of all the hard or symbolic links of the man
pages scattered throughout the filesystem, and also for clean up of the MLINKS
mess in the makefiles.
\end{description}
\subsection{Tools Used}
The two main tasks that are needed to be performed are parsing of the man pages
and building an index of the data. \textit{libmandoc} [6]
and \textit{sqlite} [7] have been used for the two tasks respectively in this
implementation.
\begin{description}
\item[\textit{libmandoc}] \hfill \\
\textit{libmandoc} is a library interface to a validating compiler. It provides
interface to parse and build an AST (Abstract Syntax Tree) of the man page, 
and interface for traversing that tree in order to extract the data from nodes
which are of our interest.
\end{description}
\begin{description}
\item[\textit{sqlite}] \hfill \\
\textit{sqlite} is an embedded relational database management system providing a
relatively small and easy to use C library interface. One of the main reason
for choosing it over the myriad of other possible options is that it provides
in built support for full text search through it's FTS virtual table module [8].
The FTS module can be accessed using pretty much standard SQL syntax, and it is
still flexible enough to accept user supplied ranking function to suit the needs
of the application. Besides that another advantage of \textit{sqlite} is the
RDBMS support which makes it very easy to store additional metadata in the form
of normal database tables without any hassles.
\end{description}
\section{Implementation Details}
Due to space constraints it is not possibile to go into enough implementation details but in this section some important components of this project are
discussed in brief.
\begin{description}
\item[makemandb] \hfill \\
\textit{makemandb} [9] is the key component of this implementation. It is a
command line tool which
traverses the filesystem, reads the raw man page source files, feeds them to the
\textit{libmandoc} parser and then stores the extracted data in the database
using \textit{sqlite}.
\end{description}
\begin{description}
\item[apropos] \hfill \\
This is the version of \textit{apropos} written from scratch utilizing the
full text search functionality of \textit{sqlite}. Unlike the classical version
of \textit{apropos}, it only displays the top 10 results relevant to the user
query and most of the times this is sufficient as will be seen later. Also, it
shows a brief snippet for each of the search results, making it more easy to
identify the relevant documents.
\end{description}
\begin{description}
\item[apropos-utils] \hfill \\
\textit{apropos-utils} [10] is a small library interface provided with this
implementation. It provides functions for querying the FTS index and for
processing the results in a user supplied callback function. It's main purpose
is to develop different interfaces on top of this implementation for different
use cases. For example a small CGI application was built using it for doing
the searches from a web browser, similarly an IRC bot was also developed utilizing this interface.
\end{description}
\section{Results}
This section shows results of some of the sample queries on the version of
developed as part of this project
\textit{apropos} to demonstrate how it solves many of the problems cited earlier
with the classical version of \textit{apropos}.
\footnote{Although this implementation of \textit{apropos(1)} returns 10 results
in normal case for a query but in the following listings the output has been
snipped to save space
}
\lstset{language = sh, caption={Add new user},}
\begin{lstlisting}
$ apropos ``add new user''
ssh-add(1)      adds private key identities to the authentication agent
...on the command line. If any file requires a passphrase, ssh-add
asks for the passphrase from the user. The passphrase is read from the
user's tty. ssh-add retries the last passphrase if multiple identity
files are given...

@@

mkdir(2)        make a directory file
...will contain the directory has been exhausted. EDQUOT The user's
quota of inodes on the file system on which the directory is being
created has been exhausted. EIO An I/O error occurred while making the
directory entry or...
\end{lstlisting}
\lstset{language = sh, caption={signal number to string},}
\begin{lstlisting}
$ apropos ``signal number to string''
psignal(3)      system signal messages
...the signal number is not recognized sigaction(2) , the string
Unknown signal is produced. The psiginfo function produces the same
output as the psignal function, only it uses the signal number
information from the si argument. The message strings can...

intro(2)        introduction to system calls and error numbers
...undefined signal to a signal(3) or kill(2) function). 23 ENFILE Too
many open files in system . Maximum number...shell. Pathname A path
name is a NUL -terminated character string starting with an optional
slash \&/ , followed by zero or...

groff_mdoc(7)   reference for groff's mdoc implementation
...Qq .Qq string ) , string ) , .Qq string Ns ), string ), .Sq .Sq
string string .Em or...UNTITLED is used. The section number may be a
number in the range 1...2, 3 and 9 error \&.\e"     and signal
handling only. \&.\e" .Sh ERRORS \&.\e...
\end{lstlisting}

\begin{figure}[htp]
\includegraphics[scale=0.30]{/home/abhinav/development/AsiaBSDCon/spell.png}
\caption{The spell corrector and the web interface in action}
\label{}
\end{figure}
\section{Related Work}
There are at least two projects which are related to this in some way.
\begin{description}
\item[man-db] \hfill \\
\textit{man-db} [11] is a complete implementation of the man page documentation
system and it is used on a number of GNU/Linux distributions.\textit{man-db}
takes an interesting approach for indexing the man page data. Unlike the
classical \textit{apropos}, it uses a Berkley DB database but still it's index
is limited to the NAME section only. It adds an option `K' to man(1) to allow a
crude full text search but it is not very efficient nor effective.
\end{description}
\begin{description}
\item[mandocdb] \hfill \\
\textit{mandocdb} [12] is more in line with the goals of our project but it takes
a novel approach. It indexes the keywords extracted from the man pages in a
key-value store using btree(3) [13]. It also comes with it's own implementation
of \textit{apropos} which performs search using this key-value store. The
main key point of this implementation is that it exploits the semantic structure
of the man pages. 
\end{description}
\section{Future Work}
\begin{description}
\item[Work on Ranking Algorithm] \hfill \\
A ranking algorithm based on probabilistic model [14] of information
retrieval has been implemented to filter out the most relevant results and show
them at the top. It is essentially based on the Okapi BM25F algorithm [15] and
uses certain parameters whose values are usually dependent on the corpus and the
search application. For example certain weight parameters are assigned to different sections of man pages. At the moment these values have been determined
manually but one goal is to use some machine learning techniques in the coming
days to automate this.
\end{description}
\begin{description}
\item[Fix Some Loose Ends In Parsing of man pages] \hfill \\
Although the parsing routine is working fine for most purposes on the man pages
supplied with NetBSD and also on most of the man pages found in Pkgsrc. But there
are perhaps a few corner cases still to be fixed. Besides that another issue that
is to be addressed is the parsing of the escape sequences which are used
extensively in the \textit{man(7)} [16] based man pages. These issues are to be
addressed in the coming days.
\end{description}
\section{Acknowledgement}
This project has been developed as part of \textbf{Google Summer of Code 2011}
[17], so thanks to \textbf{Google} for sponsoring it. Special thanks to
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. We would like to thank \textbf{David Young} who was involved with
this project closely and offered useful help and guidance throughout. A special
thanks goes to \textbf{Thomas Klausner} who helped in writing and reviewing our
man pages. Thanks to \textbf{Petra Zeidler} for administering the GSoC program
for \textit{The NetBSD Foundation}.
\begin{thebibliography}{20}
\bibitem{1}
\emph{NetBSD manual page for apropos(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?apropos++NetBSD-5.1
